Meta-learning in movement prediction problem of aperiodic time-series data

### 2025-03-03 doc/thesis/chapters/4.experiment.tex
paper + thesis

Foreign exchange in particular and financial indices in general are typical data types for aperiodic time-series data. Therefore, we choose this type of data to test the model. Specifically, we configure two datasets using foreign exchange data. \verb|USD/JPY| dataset consists of only data of the exchange rate between US dollar and Japanese yen. The data is sampled hourly from 2000 to 2024, including the attributes of open, low, high, and close price. \textbf{By carefully examining the transaction history as well as effectively aggregating data characteristics, we expect to forecast the movement of price without utilizing external information.}

@@

We evaluate each component in \verb|Temporal-ML| as well as the choice of temporal feature extraction method by removing or replacing them from the algorithm and compare the results with the original algorithm. Specifically, we conduct experiments on two groups of algorithms: (1) - Model without \verb|BiLSTM|; (2) - Model without ML. Regarding the evaluation of these models, ML-based models are evaluated as \verb|Temporal-ML|, the remaining models are evaluated as \verb|NHITS|.

>>>>

Foreign exchange in particular and financial indices in general are typical data types for aperiodic time-series data. Therefore, we choose this type of data to test the model. Specifically, we configure two datasets using foreign exchange data. \verb|USD/JPY| dataset consists of only data of the exchange rate between US dollar and Japanese yen. The data is sampled hourly from 2000 to 2024, including 4 attributes: \textit{open, low, high, and close price}. \textbf{By carefully examining the transaction history as well as effectively aggregating data characteristics, we expect to forecast the movement of price without utilizing external information.}

@@

We evaluate each component in \verb|Temporal-ML| by removing or replacing it from the algorithm and compare the results to the original algorithm. Specifically, we conduct experiments on two groups of algorithms: (1) - Model without \verb|BiLSTM|; (2) - Model without ML. Regarding the evaluation of these models, ML-based models are evaluated as \verb|Temporal-ML|, the remaining models are evaluated as \verb|NHITS|.

### 2025-03-03 doc/thesis/chapters/5.results.tex
paper + thesis

\begin{table}[h]
    \centering
    \cprotect\caption{Accuracy (\%) of \verb|Temporal-ML| and \verb|NHITS| on each attribute of \verb|USD/JPY| and \verb|multi-fx|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
    \label{tab:mine_nhits_att}
    \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|cccc}
    \toprule
    \multicolumn{1}{c}{}                &                                & $Open$                  & $High$                  & $Low$                   & $Close$                     \\
    \hline
    \multirow{3}{*}{\Verb|USD/JPY|}     & Random                         & $49.91$                 & $49.91$                 & $50.33$                 & $50.66$                     \\
                                        & \Verb|NHITS|                   & $75.03$                 & $53.74$                 & $51.57$                 & $53.50$                     \\
                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{93.06\pm1.36}$ & $\mathbf{68.87\pm1.53}$ & $\mathbf{51.01\pm1.58}$ & $\mathbf{68.37\pm2.16}$     \\
    \hline
    \multirow{3}{*}{\Verb|multi-fx|}    & Random                         & $49.92\pm2.19$          & $49.93\pm2.22$          & $49.94\pm2.17$          & $49.85\pm2.13$              \\
                                        & \Verb|NHITS|                   & $58.36\pm7.13$          & $50.66\pm4.28$          & $51.17\pm3.21$          & $53.84\pm4.60$              \\
                                        & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{78.13\pm6.33}$ & $\mathbf{62.74\pm3.30}$ & $\mathbf{58.28\pm3.86}$ & $\mathbf{65.88\pm12.49}$    \\
    \bottomrule
    \end{tabular}}
\end{table}

@@

\begin{table}[h]
    \centering
    \cprotect\caption{Classification results (\%) of \verb|Temporal-ML| and models without \verb|BiLSTM|. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
    \label{tab:ab_lstm}
    \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|cccc} 
    \toprule
    \multicolumn{1}{c}{}      &                                          & $accuracy$               & $precision$              & $recall$                 & $F1-score$                \\ 
    \hline
    \multirow{3}{*}{\Verb|USD/JPY|} & \Verb|Temporal-ML|\mbox{*}         & $\mathbf{70.33\pm1.69}$  & $\mathbf{70.73\pm1.82}$  & $\mathbf{70.26\pm1.93}$  & $\mathbf{69.28\pm2.50}$   \\
                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $68.75\pm1.78$           & $69.15\pm1.55$           & $68.73\pm1.77$           & $69.14\pm2.17$            \\
                                    & \Verb|MAML|(\Verb|Transformers|)   & $60.67\pm4.09$           & $59.74\pm6.24$           & $59.32\pm3.79$           & $53.15\pm6.19$            \\ 
    \hline
    \multirow{3}{*}{\Verb|multi-fx|}& \Verb|Temporal-ML|\mbox{*}         & $\mathbf{66.26\pm7.45}$  & $\mathbf{67.08\pm8.97}$  & $\mathbf{65.76\pm7.79}$  & $\mathbf{64.06\pm10.00}$  \\
                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $66.08\pm8.19$           & $66.56\pm9.23$           & $65.41\pm8.66$           & $63.78\pm10.83$           \\
                                    & \Verb|MAML|(\Verb|Transformers|)   & $56.03\pm6.08$           & $56.32\pm6.01$           & $55.63\pm5.01$           & $52.33\pm7.08$            \\ 
    \hline
    \multirow{3}{*}{\Verb|ETT-m2|}  & \Verb|Temporal-ML|\mbox{*}         & $71.14\pm9.13$           & $62.21\pm7.51$           & $58.75\pm5.30$           & $57.61\pm6.74$            \\
                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $\mathbf{71.40\pm9.10}$  & $\mathbf{63.25\pm8.82}$  & $\mathbf{60.22\pm8.42}$  & $\mathbf{59.43\pm9.24}$   \\
                                    & \Verb|MAML|(\Verb|Transformers|)   & $68.69\pm8.85$           & $58.54\pm4.38$           & $57.85\pm4.71$           & $57.07\pm5.14$            \\ 
    \hline
    \multirow{3}{*}{\Verb|WTH|}     & \Verb|Temporal-ML|\mbox{*}         & $\mathbf{74.97\pm2.63}$  & $\mathbf{69.13\pm3.44}$  & $\mathbf{65.98\pm3.96}$  & $\mathbf{66.00\pm3.80}$   \\
                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $74.41\pm2.71$           & $68.17\pm4.60$           & $65.50\pm4.82$           & $65.63\pm4.47$            \\
                                    & \Verb|MAML|(\Verb|Transformers|)   & $72.28\pm3.06$           & $62.35\pm8.73$           & $60.90\pm4.97$           & $58.65\pm4.72$            \\
    \bottomrule
\end{tabular}}
\end{table}

>>>>

\begin{table}[h]
    \centering
    \cprotect\caption{Accuracy (\%) of \verb|NHITS| on each attribute of \verb|USD/JPY| and \verb|multi-fx| dataset.}
    \label{tab:mine_nhits_att}
    % \resizebox{\columnwidth}{!}{}
    \begin{tabular}{c|c|cccc}
        \toprule
        \multicolumn{1}{c}{}                &                                & $Open$                  & $High$                  & $Low$                   & $Close$                     \\
        \hline
        \multirow{2}{*}{\Verb|USD/JPY|}     & Random                         & $49.91$                 & $49.91$                 & $50.33$                 & $50.66$                     \\
                                            & \Verb|NHITS|                   & $75.03$                 & $53.74$                 & $51.57$                 & $53.50$                     \\
                                            % & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{93.06\pm1.36}$ & $\mathbf{68.87\pm1.53}$ & $\mathbf{51.01\pm1.58}$ & $\mathbf{68.37\pm2.16}$     \\
        \hline
        \multirow{2}{*}{\Verb|multi-fx|}    & Random                         & $49.92\pm2.19$          & $49.93\pm2.22$          & $49.94\pm2.17$          & $49.85\pm2.13$              \\
                                            & \Verb|NHITS|                   & $58.36\pm7.13$          & $50.66\pm4.28$          & $51.17\pm3.21$          & $53.84\pm4.60$              \\
                                            % & \Verb|Temporal-ML|\mbox{*}     & $\mathbf{78.13\pm6.33}$ & $\mathbf{62.74\pm3.30}$ & $\mathbf{58.28\pm3.86}$ & $\mathbf{65.88\pm12.49}$    \\
        \bottomrule
        \end{tabular}
\end{table}

@@

\begin{table}[h]
    \centering
    \cprotect\caption{Classification results (\%) of \verb|Temporal-ML| and models with \verb|BiLSTM| replaced. Best results per metrics are boldfaced. (\mbox{*}): Our method.}
    \label{tab:ab_lstm}
    \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|cccc} 
    \toprule
    \multicolumn{1}{c}{}      &                                          & $accuracy$               & $precision$              & $recall$                 & $F1-score$                \\ 
    \hline
    \multirow{3}{*}{\Verb|USD/JPY|} & \Verb|Temporal-ML|\mbox{*}         & $\mathbf{70.33\pm1.69}$  & $\mathbf{70.73\pm1.82}$  & $\mathbf{70.26\pm1.93}$  & $\mathbf{69.28\pm2.50}$   \\
                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $68.75\pm1.78$           & $69.15\pm1.55$           & $68.73\pm1.77$           & $69.14\pm2.17$            \\
                                    & \Verb|MAML|(\Verb|Transformers|)   & $60.67\pm4.09$           & $59.74\pm6.24$           & $59.32\pm3.79$           & $53.15\pm6.19$            \\ 
    \hline
    \multirow{3}{*}{\Verb|multi-fx|}& \Verb|Temporal-ML|\mbox{*}         & $\mathbf{66.26\pm7.45}$  & $\mathbf{67.08\pm8.97}$  & $\mathbf{65.76\pm7.79}$  & $\mathbf{64.06\pm10.00}$  \\
                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $66.08\pm8.19$           & $66.56\pm9.23$           & $65.41\pm8.66$           & $63.78\pm10.83$           \\
                                    & \Verb|MAML|(\Verb|Transformers|)   & $56.03\pm6.08$           & $56.32\pm6.01$           & $55.63\pm5.01$           & $52.33\pm7.08$            \\ 
    \hline
    \multirow{3}{*}{\Verb|ETT-m2|}  & \Verb|Temporal-ML|\mbox{*}         & $71.14\pm9.13$           & $62.21\pm7.51$           & $58.75\pm5.30$           & $57.61\pm6.74$            \\
                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $\mathbf{71.40\pm9.10}$  & $\mathbf{63.25\pm8.82}$  & $\mathbf{60.22\pm8.42}$  & $\mathbf{59.43\pm9.24}$   \\
                                    & \Verb|MAML|(\Verb|Transformers|)   & $68.69\pm8.85$           & $58.54\pm4.38$           & $57.85\pm4.71$           & $57.07\pm5.14$            \\ 
    \hline
    \multirow{3}{*}{\Verb|WTH|}     & \Verb|Temporal-ML|\mbox{*}         & $\mathbf{74.97\pm2.63}$  & $\mathbf{69.13\pm3.44}$  & $\mathbf{65.98\pm3.96}$  & $\mathbf{66.00\pm3.80}$   \\
                                    & \Verb|MAML|(\Verb|BiLSTM+CNN|)     & $74.41\pm2.71$           & $68.17\pm4.60$           & $65.50\pm4.82$           & $65.63\pm4.47$            \\
                                    & \Verb|MAML|(\Verb|Transformers|)   & $72.28\pm3.06$           & $62.35\pm8.73$           & $60.90\pm4.97$           & $58.65\pm4.72$            \\
    \bottomrule
\end{tabular}}
\end{table}
